OvRF review

### 2020-09-14 paper/main.tex
Working on methods section

\section{Materials and Methods}
\subsection{Overview of overlapping reading frames}
Initially, we downloaded an accession list of all viral genomes from NCBI \url{https://www.ncbi.nlm.nih.gov/genome/viruses/} that has information regarding: representative, neighbor, host, taxonomy and segment name for every virus entry on the database. 
% Re-making this part of the analysis. Will have to updated scripts and methods
We then used a Python script to complement information of the entries to include taxonomical information, genome lengt and total number of proteins. 
The same script was used to download all of the proteins associated with each accession number and their position on the genome. 
We used the protein file to find overlapping proteins and calculate the frameshift and number of nucleotides involved in each overlap/ 
We also created a script to classify each viral family according to the Baltimore classification displayed on ViralZone \url{https://viralzone.expasy.org/}.
The final result is a data base with a total of 11,891 viral entries, from which 9,928 are classified in any Baltimore groups. 
From remaining 1,963 entries, we can say, according to the information obtained from NCBI, that 400 are DNA viruses and 1,552 are RNA.

\subsection{Adjacency graphs}
To analyze the organization and distribution of overlapping reading frames in different virus families, we downloaded for each family all the reference complete genomes from \url{https://www.ncbi.nlm.nih.gov/labs/virus/vssi/#/virus?SeqType_s=Nucleotide&SourceDB_s=RefSeq} as a fasta file. 
Over that file, we run a Python script to measure kmer-distance to create a distance matrix.
We then used that matrix as an input to form clusters representing similar proteins. 
Once we have each protein classified in a specific cluster, we generate a network plot, where each node represents a cluster, and the edges between clusters are formed when two proteins are adjacent to each other.

>>>>

\section{Materials and Methods}
\subsection{Data collection and processing}
The initial part of the analysis consisted on the creation of the database.
A list of all virus genomes was downloaded from NCBI \url{https://www.ncbi.nlm.nih.gov/genome/viruses/}.
The table had information regarding: representative, neighbor, host, taxonomy and segment name for every reference genome on the database. 
The neighbor information is a new approach from NCBI designed to include more diversity for every species, so every reference genome have some other genomes associated to it that are also complete or near complete. 
However, for the purpose of this study, we considered that the reference genome of each species is sufficient information. 
Information of only the reference genomes was recorded and complemented by pulling data from the GeneBank database including: genome length, number of proteins, topology and molecule type using a Python script called scraper-2020.py.
The same script was used to create a different table in which we stored all the CDSs of every genome with accession number (of the genome they belong to), product, strand, coordinates and start codon position. 
A second script (find$_$ovrfs.py) was used to group rows of the CDSs table by accession number and compare the coordinates of each protein. The output is a table with accession number of the reference genome, product, location and length of the overlapping proteins, number of nucleotides involved in the overlap and frame shift.
%Update this information with new downloads
We also created a script to classify each viral family according to the Baltimore classification displayed on ViralZone \url{https://viralzone.expasy.org/}.
The final result is a data base with a total of 11,891 viral entries, from which 9,928 are classified in any Baltimore groups. 
From remaining 1,963 entries, we can say, according to the information obtained from NCBI, that 400 are DNA viruses and 1,552 are RNA.

\subsection{Overview of overlapping reading frames in virus genomes}

\subsection{Clustering protein data}
To analyze the organization and distribution of overlapping reading frames in different virus families, we downloaded for each family all CDSs of the reference genomes from \url{https://www.ncbi.nlm.nih.gov/labs/virus/vssi/#/virus?SeqType_s=Nucleotide&SourceDB_s=RefSeq}.
The information for all CDSs was stored in a multi-fasta file where each entry consisted in a header with the name of the virus, the accession number of the genome, the product name of the CDS, the strand in which it is encoded, the coordinates in the genome and the aminoacid sequence.
Because aligned-based methods are computationally expensive, specially for highly divergent sequences, we used a k-mer frequency method to compare the aminoacid sequences of virus families from different Baltimore groups.
The distance matrix generated from the k-mer distance measure, is used as an input for an R script that uses a clustering method to clasify proteins in clusters according to their similarities. 
% Describe clustering method

\subsection{Creation of adjacency graphs}
A Python script was created to automatically generate adjacency plots to represent the organization of viral genomes across the family. 
The script use as an input a csv file in which each line is a protein with information regarding the genome it belongs to, the coordinates where it is located and the cluster it belongs to. 
We use the information to find the order of the proteins for each genome and labeled them depending on if they overlap or not. 
Using Graphviz (Graph Visualization Software) \url{https://graphviz.org/theory/} to create a dot file of a network graph where each node represents a cluster, and each edge represent adjacent proteins between two clusters.
Overlapping proteins are colored in blue, weather adjacent proteins are colored in gray. 
As complementary information, the script uses Matplotlib\cite{Hunter:2007} to generate maps of each viral genome (where each protein is colored according to their cluster classification), and a wordcloud to visualize the names of the proteins in each cluster.
